---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

### 1. Load the data

```{r loading, echo=TRUE}
activity <- read.csv("activity.csv")
head(activity)
```


## What is mean total number of steps taken per day?

### 1. Calculate the total number of steps taken per day

```{r total seps, echo=TRUE}
library("dplyr")
act0 <- activity %>% group_by(date) %>% summarise(total.steps = sum(steps,na.rm=TRUE))
```


### 3.  Calculate and report the mean and median 
```{r mean, echo=TRUE}
mean(act0$total.steps,na.rm=TRUE)
```

```{r median, echo=TRUE}
median(act0$total.steps,na.rm=TRUE)
```

### 2.  Histogram
```{r barplot, echo=TRUE}
barplot(act0$total.steps, xlab = "Days",ylab="Total Steps")
```


## What is the average daily activity pattern?

### 1. Make a time series plot
```{r mean steps per interval, echo=TRUE}
act1 <- activity %>% group_by(interval) %>% summarise(mean.steps = mean(steps,na.rm=TRUE))
```

```{r plot, echo=TRUE}
plot(x=act1$interval,y=act1$mean.steps,type="l",xlab = "Intervals",ylab="Steps mean")
points(835,206.170,pch=20,col="red")
```

### 2. Which 5-minute interval contains the maximum number of steps?
```{r max, echo=TRUE}
max(act1$mean.steps)
```

Max = 206.170  =>  Interval = 835


## Imputing missing values

### 1. Calculate and report the total number of missing values
```{r missing value, echo=TRUE}
head(colSums(is.na(activity)))
```

There is only missing values in steps column.


### 2/3.  Create a new dataset using mean calculated previously

```{r fill, echo=TRUE}
act2 <- activity

for (i in 1:17568) {
    if (is.na(act2[i,1])){
        val<- subset(act1,act1$interval == act2[i,3])
        act2[i,1] <- val$mean.steps
    }
}
```

### 4. Make a histogram 
```{r mean3, echo=TRUE}
act3 <- act2 %>% group_by(date) %>% summarise(total.steps = sum(steps,na.rm=TRUE))
```

```{r mean2, echo=TRUE}
mean(act3$total.steps,na.rm=TRUE)
```

```{r median2, echo=TRUE}
median(act3$total.steps,na.rm=TRUE)
```

```{r barplot2, echo=TRUE}
barplot(act3$total.steps , xlab = "Days",ylab="Total Steps")
```

The mean and the median has increased.

## Are there differences in activity patterns between weekdays and weekends?

### 1. Create a new factor variable 

```{r factor, echo=TRUE}
library(lubridate)

act2$day <- wday(act2$date)

act2$w.days <- ifelse(act2$day %in% 2:6,"weekday","weekend")
act2$w.days <- as.factor(act2$w.days)
```

### 2.Make a panel plot  

```{r panel plot, echo=TRUE}
act4 <- act2 %>% group_by(w.days,interval) %>% summarise(mean.steps = mean(steps,na.rm=TRUE))


library(lattice)
with(act4, xyplot(act4$mean.steps ~ act4$interval | act4$w.days , type = "l",xlab="Interval",ylab = "Mean steps"))
```










